Exercising choice over feedback schedules during practice is not advantageous for motor learning

### 2022-08-14 docs/revision-1/r1_response-to-reviewers.Rmd

\emph{In my opinion, the presented manuscript is well-written and contributes a well-powered study to the current debate surrounding self-controlled learning. However, I think a stronger rationale for Experiment 2 (binary feedback) should be provided. It is not clear to me why this manipulation was presented as a standalone experiment rather than as additional feedback condition in Experiment 1, and perhaps it is not necessary to include the binary feedback groups in the manuscript at all. Additionally, as the manuscript discusses both the OPTIMAL and information-processing perspectives as potential explanations for self-controlled learning advantages, I feel that the inclusion of further analyses regarding participants’ feedback requests could help to better understand the mechanisms underpinning self-controlled learning. Overall, I generally believe that this paper is publishable pending some relatively minor revisions. I have noted page and line numbers (according to the authors’ submitted version) along with my more specific points below.}}

Thank you for the positive comments about our manuscript. In line with comments from the other two Reviewers, we have updated the Introduction to more clearly explain our rationale for and the value of Experiment 2. In short, after we had analyzed the data from Experiment 1, we realized that a second experiment was required where participants received a type of feedback that would have a very low informational value for how to correct one's behaviour (i.e., would not generate an error signal). This is why binary feedback was chosen for Experiment 1 because both error and graded feedback generate an error signal as both provide information about how to change one's behaviour. As such, we feel Experiment 2 was a stronger test of the information-processing and motivational perspectives. From the information-processing view, you'd expect no performance advantages of having choice over binary feedback in any experimental phase whereas from the motivational view, you'd still expect performance advantages in practice, retention, and transfer because these benefits arise out of having choice. Additionally, keeping the results of Experiment 2 in the manuscript is important for meta-analytic reasons, especially given the recent self-controlled learning meta-analysis found that the putative self-controlled learning advantage is the result of publication bias rather than a true effect [@mckay2021].

> Introduction (Page 4 line 91 continued on page 5 lines 92-100):
>
> To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, we manipulated the amount of information participants in choice and yoked (i.e., no-choice) groups experienced with their feedback schedule during acqusition of a novel motor task. In Experiment 1, participants received error or graded feedback to assess how high and moderate levels of informational value impact the self-controlled learning advantage. Given both error and graded feedback provide salient information about how to correct one's behavior relative to the task goal (i.e., both generate an error signal), in Experiment 2 we provided participants with binary feedback. As binary feedback is devoid of information about the necessary change to improve one's behavior (i.e., does not generate an error signal), we could better isolate the motivational nature of choice to test between the two explanations for the self-controlled learning advantage.

\Done

\RC{\emph{p. 7, lines 16-17: “Additionally, half of the participants in each group were randomly selected to estimate their performance on the spatial and timing goals after each trial in the pre-test.” Please provide more detail on how participants provided their estimations (e.g., manual vs. verbal responses, spatial and/or timing estimations)}}

We have clarified that participants made their estimations verbally on pages 7 (line 168) and 11 (line 210).

\Done

\RC{\emph{p. 9, lines 1-2: “Following a 2000 ms feedback delay interval, the feedback decision prompt was presented for the self-controlled groups.” How did participants indicate whether they did or did not want feedback? (e.g., manual vs. verbal response?) Although this is nicely illustrated in Figure 1, it would be helpful to also include in the text that participants were made aware of how many feedback trials remained}}

We have clarified that participants made their feedback request verbally on page 9 (line 208) and in the Figure 1 caption (page 10). We have also included the information about the remaining feedback trials in the text on page 9 (lines 205-206).

\Done

\RC{\emph{p. 8, lines 17-18: “Graded feedback for the spatial goal was provided as “too short” if performance was < 40 degrees (or 60 degrees in transfer)”. I think it’s confusing to include the ’60 degrees in transfer’ here (as well as on page 9, lines 18-20) if there was no feedback provided in transfer (according to p. 7, lines 13-14: “No feedback about motor performance was provided in pre-test, retention, or transfer”).}}

Thank you for catching this. We have removed this confusing information.

\Done

\RC{\emph{p. 9, lines 9-12: “The perceived competence and task interest and enjoyment questions were from the Intrinsic Motivation Inventory and the perceived autonomy questions were used in earlier work”. Just to clarify, the perceived autonomy questions are not the same as the perceived choice subscale in the IMI? If this is a customized scale, please note if the specific items are available in a previous publication, and if not, consider including them in the supplementary materials.}}

The perceived autonomy questions we used are not the perceived choice subscale of the IMI. You are correct that they are a customized scale that we have used in the past [e.g., @carter2017b; @stgermain2022]. We have also added the questions we used for all questionnaires to the public repository for this project in the directory "materials".

\Done

\RC{\emph{Is the finding that exercising choice did not impact perceptions of autonomy compared to yoked groups unusual? Is there a different/perhaps more sensitive way to assess perceptions of autonomy?}}

It is actually quite common to not find an effect of choice over some practice variable like one's feedback schedule on measures of perceived autonomy [e.g., @barros2019; @carter2017b; @mckay2020b; @stemarie2013]. It was actually only very recently that some studies found higher perceptions of autonomy [e.g., @mckay2020a; @stgermain2022] in those participants given choice versus those not given the same choice opportunities. A key difference is both of those experiments had even larger sample sizes per group than what we had in the present work. For instance, St. Germain et al. (2022) had 50 participants per group. Thus, we think this highlights that the impact of being given choice over a practice variable has a rather small effect on perceived autonomy and requires larger sample sizes to reliably detect. As a side note, our perceived autonomy questions are those used in the St. Germain et al. paper, which found an effect favouring choice on perceived autonomy.

\Done

\RC{\emph{I felt there were a number of additional analyses that could be conducted to help provide a more fulsome story. For example:}}

Thank you for the following suggestions. In short, we did not perform any analyses regarding the distribution of feedback requests or how feedback requests related to performance. Both of these have been reported in several previous papers investigating self-controlled feedback schedules. We provide more details regarding our decision below when we respond to each analysis suggestion independently. We did, however, include details regarding the number of "hits" as this was something that Reviewer 2 also raised.

\Done

\RC{\emph{a. What was the distribution of feedback requests throughout acquisition? (e.g., did participants generally choose to fade their feedback? or perhaps are there early vs. late feedback requesters?)}}

Our main reason for not analyzing the distribution of feedback requests or early verus late requesters is these are not things we manipulated in our design. Thus, these analyses would be post-hoc and non-causal. As such, we do not think that would provide any new insight beyond what has already been reported in earlier self-controlled feedback research: sometimes participants seem to create faded request schedules [e.g., @chiviacowsky2002; @janelle1997] and sometimes schedules stay relatively consistent [e.g., @patterson2010]. Additionally, a recent meta-analysis [@mckay2022] did not find evidence for several feedback scheduling hypotheses: no evidence 1) that faded schedules are more effective than non-faded, 2) that performance-contingent or bandwidth schedules are more effective than fixed schedules, and 3) that feedback frequency does not have consistent effects. For these reasons, we elected to not conduct these analyses and instead keep our discussion more focused on the fact that our results are consistent with the recent self-controlled learning meta-analysis suggesting this supposed effect is likely non-existent and instead was the result of publication bias.

\Done

\RC{\emph{b. How did feedback requests relate to learners’ performance? (e.g., did feedback requests tend to follow relatively good vs. bad trials?). This seems particularly important to address with reference to the information-processing explanation of self-controlled learning advantages.}}

We did not query participants on their reasons why or why not feedback was requested during practice. Without knowing this information for each participant, it is not possible to know whether any relation to relatively good or relatively bad trials was strategic or coincidental. Given such an analysis would also be post-hoc and non-causal, we also do not think this analysis would provide new insight beyond that already reported in earlier self-controlled feedback experiments that have used multiple-choice questionnaires [e.g., @carter2012; @chiviacowsky2002; @patterson2010; @patterson2011] or open-ended questions [e.g., @carter2016; @laughlin2015]. Similar to above, not including these unplanned analyses allows us to remain focused on our main conclusion that there is a lot of evidence emerging that strongly suggests the supposed self-controlled learning advantage is not a replicable or real effect.

\Done

\RC{\emph{c. How often did participants actually "hit" the target? It seems that the success criteria were quite strict - can you be expected to “acquire/retain” a skill you haven’t successfully performed? Could the potential unlikelihood of success/limited expectancies for success have undermined potential motivational aspects of self-control?}}

We have added information about the number of "hits" in Table 2 (Page 18) of the revised manuscript based on your comment and that from Reviewer 2. Interestingly, the "hit" rate was low in both experiments and for all groups. Although the low success rates seem to have impacted perceptions of competence, particularly for those that received graded and binary feedback, the graded feedback groups improved their performance during practice, and retention and transfer were better than pre-test levels. Further, the graded feedback groups' ratings for perceived competence were numerically similar to those reported by the participants in Experiment 2 with binary feedback. Additionally, self-reported intrinsic motivation was numerically similar in all experimental groups. As such, we do not think the low "hit" rates in the present experiments undermined any potential motivational influences of self-controlled feedback schedules. We also now make reference to the hit data in the Discussion.

> Discussion (Page 23 Lines 447-450):
>
> The number of "hits" for the spatial and timing goals were both quite low for all groups. Although this may have also impacted perceptions of competence, the relatively low "hit" rate did not seem to differentially impact intrinsic motivation as self-reported levels were quite similar for all groups.

\Done

\RC{\emph{p. 18, lines 1-3: “...argued that self-controlled feedback is effective because it provides the opportunity to request feedback in a performance dependent way that reduces uncertainty about movement outcomes relative to task goals”. Yes – as noted above, it would be helpful to understand whether these participants requested feedback in a performance dependent fashion}}

In line with our response above, we did not query participants on their reasons for requesting feedback. The idea that self-controlled feedback is effective because it reduces uncertainty about movement outcomes emerged from early work by @chiviacowsky2005 that showed a self-controlled feedback schedule was only effective if the feedback decision was made after and not before one's performance. Carter et al. (2014) extended this work with a group that made their feedback decision before a trial but could then change their mind after a trial. Thus, the key idea behind requesting feedback in a performance dependent way to reduce uncertainty is about making the decision *after* one's movement, which was the case for all self-controlled groups in the present experiment. But consistent with earlier responses and our stance in the manuscript, we do not think the self-controlled learning advantage is a real effect based on recent experiments with large samples (and often pre-registered) and the meta-analysis by @mckay2021. As such, analyses and discussions surrounding the information-processing explanation or the motivational (i.e., OPTIMAL theory) view seem moot. We instead contend that contextualizing our results within these larger, and often pre-registered, experiments and the McKay et al. meta-analysis is the more critical discussion point.

\Done

\RC{\emph{p. 19, lines 19-21: “Similarly, self-controlled feedback schedules did not enhance error estimation skills (see Supplementary A) [...], inconsistent with the information-processing perspective.” Similar to above, I think we need to know more about participants feedback choices to consider why error estimation skills did not improve and how this speaks to the information-processing perspective}}

To clarify, error estimation skills did improve from pre-test to retention and transfer for the subset of participants who estimated during pre-test in Experiment 1. We did not find that error estimation was enhanced in the choice groups compared to the yoked groups. We have made this more clear in the revised manuscript.

> Discussion (Page 22 Lines 420-421)
>
> Similarly, self-controlled feedback schedules did not enhance error estimation skills compared to yoked schedules (see **Supplementary B**) and choice did not interact with feedback characteristics, inconsistent with the information-processing perspective.

Similar to our above comments, such a finding is consistent with the idea that there is no self-controlled learning advantage to explain from an information-processing or motivational view in motor learning [@mckay2021].

\Done

\RC{\emph{p. 20, lines 6-10: “Additionally, we used a strict criteria with binary feedback where any outcome other than zero error was considered a miss. Thus, binary feedback may be more effective when paired with a tolerance zone such as that used in the bandwidth technique” Agreed... As mentioned earlier, please indicate what the rate of success was in this group and consider whether this group can be integrated into Experiment 1 or removed}}

As mentioned above with respect to your earlier comment, we have added this information in the revised manuscript in Table 2 on page 18.

\Done

# References

::: {#refs custom-style="Bibliography"}
:::
